Algorithm explanation

### 2013-08-22 article/algo.tex
Pseudo-code updated

For simplicity of discussion, we define \emph{term arity} $k_v$ for
every Horn term vertex $v \in V_T$ with incoming edges.
\begin{align*}
k_v =
\begin{cases}
\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
\end{cases}
\end{align*}
We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.

A variable mapping label on edges \theta is used to map the index of
one predicate variable to another. (TODO?)

@@

\begin{align*}
\hat C_{a,v} = &
 \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
 \bigcup_{0 < i \leq K}
 \left\lbrace
  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
 \right\rbrace \\
 & \hspace{2cm} \cup
 \left\lbrace
  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
 \right\rbrace \cup
 \bigcup_{k_v < i \leq K}
 \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
\\
C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
\begin{cases}
\emptyset
& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
 b_v = \lambda_v b_0
\right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
\end{cases}
\end{align*}

For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
from $C_u$.

@@

Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.

\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
\newcommand{\splitGraph}{\ensuremath{\mbox{\sc RestructureGraph}}}

\begin{algorithm}
\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
\begin{algorithmic}
\STATE {$v_\bot \gets $ the root node of $G$}
\STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
\REPEAT
  \STATE {$C \gets \genConstr (v_\bot, V)$}
  \STATE {$c_{v_\bot} \gets C(v_\bot)$}
  \STATE {Evaluate satisfiability of $c_{v_\bot}$}
  \IF {$c_{v_\bot}$ is unsatisfiable}
    \STATE {$\mathcal{UC} \gets $ unsatisfiable core for $c_{v_\bot}$}
    \STATE {$G \gets \splitGraph (G, \mathcal{UC})$}
    \STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
  \ELSE
    \STATE {Enlarge $V$}
    \STATE {$m \gets $ one model of $c_{v_\bot}$}
  \ENDIF
\UNTIL {$V$ contains all predicate variable vertices}
\STATE {Build solution by $m$}
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$ \genConstr (v, visited) $}\label{alg:genConstr}
\begin{algorithmic}
\STATE {$m \gets \emptyset$}
\FORALL {$v_\Rightarrow \in preds(v)$}
  \FORALL {$v_p \in preds(v_\Rightarrow)$}
    \STATE { $ l := \varphi (v_p) $ }
    \IF {$l$ is a predicate variable}
      \STATE {$m' \gets \genConstr (v_p) $}
      \STATE {$m \gets m \cup m'$}
    \ELSE
      \STATE { $id$ }
    \ENDIF
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

\subsection{Disjunction-free problems}

If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for
$G$. We now try to give a solution by relaxing the constraint
$C_{v_\bot}$.

We consider below a Horn graph whose Horn term vertices has at
most one successor.

\subsubsection{Quantifier elimination}

A set of eliminating quantifiers are obtained as $quant(n)$.

\begin{align*}
in(n) & = \bigcup_{p\in pred(n)} out(p) \\
out(n) & = gen(n) \cup in(n) \\
dup(n) & = \left\lbrace t \middle| p,q \in pred(n); p \ne q \wedge
 t \in out(p) \wedge t \in out(q) \right\rbrace \\
quant(n) & = gen(n) \cup dup(n) \cup
 (out(n) \cap \bigcup_{s \in succ(n)} quant(s))
\end{align*}

@@

\begin{align*}
\hat C_a = &
 \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
 \bigcup_{0 < i \leq K}
 \left\lbrace
  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
 \right\rbrace \\
 & \hspace{2cm} \cup
 \left\lbrace
  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
 \right\rbrace \cup
 \bigcup_{k_v < i \leq K}
 \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
\\
C_v = & \bigcup_{\edge{a}{v}} \hat C_a \cup
\begin{cases}
\emptyset
& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_{v,\emptyset} = \lambda_v \mathbf{a}_0,
 b_{v,\emptyset} = \lambda_v b_0
\right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
\end{cases}
\end{align*}

---

\begin{align*}
C_v = & \bigcup_{(u,U,\theta,C_u) \in S_v^-} C_u \cup
\bigcup_{(u,U,\theta,C_u) \in S_v^+} C_u
\\
& \cup
\left\lbrace
 \mathbf{a}_{v,\emptyset,i} =
  \sum_{(u,U,\theta,C_u) \in S_v^-} \mathbf{a}_{u,U,\theta^{-1}(i)} +
  \sum_{(u,U,\theta,C_u) \in S_v^+} \mathbf{a}_{u,U, \theta^{-1}(i)}
\right\rbrace \\
& \cup \left\lbrace
 b_{v,\emptyset} \geq
  \sum_{(u,U,\theta,C_u) \in S_v^-} b_{u,U} +
  \sum_{(u,U,\theta,C_u) \in S_v^+} b_{u,U}
\right\rbrace \\
& \cup \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,\emptyset,i} = 0 \right\rbrace
\end{align*}
such that
\begin{align*}
& S_v^- = \left\lbrace (u, U, \theta, C_u) \middle|
  u \mathop{\leadsto}^\theta v \wedge
  \left( \nexists u' \in V'; \xi (u') = u \right) \wedge
  \Delta_{V'} (u) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right\rbrace
\\
& S_v^+ = \left\lbrace (u, U, \theta, C_u) \middle|
  u \mathop{\leadsto}^\theta v \wedge
  \left( \exists u' \in V'; \xi (u') = u \wedge
  \hat \Delta (u', \edge{u}{v}) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right) \right\rbrace
\\
\end{align*}

@@

\subsubsection{Refinement from Unsatisfiable Core}

We first start the process from the initial Copy tree:
\begin{align*}
V_C = & \left\lbrace v_{v_\bot} \right\rbrace \\
E_C = & \emptyset
\end{align*}

>>>>

For simplicity of discussion, we define \emph{term arity} $k_v$ for
every Horn term vertex $v \in V_T$ with incoming edges.
\begin{align*}
k_v =
\begin{cases}
\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
0 & \mbox{if } \varphi(v) = \bot
\end{cases}
\end{align*}
We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.

A variable mapping label $\theta$ on edges is used to map the index of
one predicate variable to another. (TODO?)

@@

\begin{align*}
\hat C_{a,v} = &
 \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
 \bigcup_{0 < i \leq K}
 \left\lbrace
  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
 \right\rbrace \\
 & \hspace{2cm} \cup
 \left\lbrace
  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
 \right\rbrace \cup
 \bigcup_{k_v < i \leq K}
 \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
\\
C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
\begin{cases}
\emptyset
& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
 b_v = \lambda_v b_0
\right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
\end{cases} \\
(TODO???)
\end{align*}
For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
from $C_u$.

@@

If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
We now try to give a less simpler solution by transforming $G$ and
relaxing the constraint $C_{v_\bot}$.  This transformation takes place
by considering the unsatisfiable core from the constraint.

Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our Horn clause
solving algorithm.  It takes an input Horn graph and returns a mapping
between predicate variables and linear arithmetic formulas.

\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}

\begin{algorithm}
\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
\begin{algorithmic}
\STATE {$v_\bot \gets $ the root node of $G$}
\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
\REPEAT
  \STATE {$\mathcal{C} \gets \genConstr (G, v_\bot, V_\star)$}
  \STATE {$C_{v_\bot} \gets \mathcal{C} \left[ v_\bot \right]$}
  \STATE {Evaluate satisfiability of $C_{v_\bot}$}
  \IF {$C_{v_\bot}$ is unsatisfiable}
    \STATE {$U \gets $ unsatisfiable core for $C_{v_\bot}$}
    \STATE {$G \gets \transGraph (G, U)$}
    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
  \ELSE
    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
m    \STATE {$m \gets $ one model of $C_{v_\bot}$}
  \ENDIF
\UNTIL {$V_\star$ contains all predicate variable vertices}
\STATE {Build solution by $m$}
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$ \genConstr (G, v, visited) $}\label{alg:genConstr}
\begin{algorithmic}
\STATE {$\mathcal{C} \gets \emptyset$}
\FORALL {$v_\Rightarrow \in preds_G(v)$}
  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
    \IF {$\varphi (v_p)$ is a predicate variable}
      \STATE {$\mathcal{C} \gets \mathcal{C} \cup \genConstr (G, v_p, visited)$}
    \ELSE
      \STATE { $id$ }
    \ENDIF
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

\subsection{Quantifier elimination}

@@

@@

\subsection{Refinement from Unsatisfiable Core}
